Hydra battles in Coq

### 2020-09-18 doc/chapter-hydras.tex
modification in *.tex

\index{To do}

\begin{todo}
Look for an existing library on trees with nodes of arbitrary arity, iin order to replace  this ad-hoc type with something more generic.
\end{todo}

@@

\subsubsection{Lexicographic order on \texorpdfstring{nat*nat}{nat*nat}}
\label{omega2-case}

We  prove now that even the type \texttt{nat * nat}, provided with the lexicographic product of \texttt{(nat,<)} by itself is too simple for proving the termination of all hydra battles. This impossibility result will prevent us from considering measures like the following one:

  \begin{Coqsrc}
  Let m h = (height h, hsize h).
  \end{Coqsrc}

The proof we are going to develop has exactly the same structure as in Section~\ref{omega-case}.
 Nevertheless, the proof of technical  lemmas is a little more complex, due to 
 the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
Consider for instance that there exists an infinite number of pairs between
$(1,0)$ and $(2,0)$.

\begin{remark}
  The order structure we consider in this section is also known as the ordinal
  $\omega^2$. We identify any pair $(i,j)\in \mathbb{N}\times\mathbb{N}$ with the ordinal $\omega\times i + j$. Thus the three kinds of ordinals in $\omega^2$
  are represented as follows:
  \begin{description}
  \item[null ordinal] : the pair $(0,0)$
  \item[successor ordinal] : any pair $(i,j)$ where $j>0$
  \item[limit ordinal] : any pair $(i,0)$ where $i>0$
  \end{description}
\end{remark}

The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.

\subsubsection{Preliminaries}
Let us assume there is a variant from \texttt{Hydra} into \texttt{nat*nat} (with the
   lexicographic ordering)  for proving the   termination of all hydra battles.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
  Section Impossibility_Proof.

  Let t := (nat * nat)%type.

  (** non-dependent lexicographic strict ordering on nat*nat *)

  Let lt2 : relation t := lexico Peano.lt Peano.lt.

  Infix "o<" := lt2.

  (** reflexive closure of lt2 *)
  Let le2 := clos_refl _ lt2.
  Infix "o<=" := le2.

  Variable m : Hydra -> t.

  Context (Hvar : Hvariant lt2_wf free m).
\end{Coqsrc}

\label{remark-infix}
\begin{remark}
We use in \coq{} scripts the infix notations \verb@o<@ and \verb@o<=@ for
the orders $<$ and $\leq$ on ordinals. This convention allows us to define
implicit coercions from \texttt{nat} into various types for representing ordinals.
For instance a proposition like \verb@ 6 o< 8@ will be interpreted as 
``the finite ordinal 6 is strictly less than the finite ordinal 8'', whereas 
\verb@ 6 o< 8@ is a statement about natural numbers. Since many lemmas talk about both ordinals and natural numbers, we wished to avoid too many scope changes inside propositions.
\end{remark}

Let us follow the same pattern as in Sect.~\ref{omega-case}.
First, we define an injection from type \texttt{t} into \texttt{Hydra}, by
 associating to  any pair $(i,j)$ the hydra with $i$ branches of length $2$ and
$j$ branches of length $1$.

%% revenir ici

\vspace{4pt}
\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
  Let iota (p: t) := 
    node (hcons_mult (hyd1 head) (fst p)
                     (hcons_mult head (snd p) hnil)).
\end{Coqsrc}

For instance, Figure~\vref{fig:essai2} shows the hydra associated to the pair
$(3,5)$.

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.4]
\node (foot) at (6,0) {$\bullet$};
\node (N1) at (1,3) {$\bullet$};
\node (N2) at (3,3) {$\bullet$};
\node (N3) at (5,3) {$\bullet$};
\node (N4) at (8,3) {$\Smiley[2][green]$};
\node (N5) at (11,3) {$\Smiley[2][green]$};
\node (N6) at (14,3) {$\Smiley[2][green]$};
\node (N7) at (17,3){$\Smiley[2][green]$};
\node (N8) at (20,3){$\Smiley[2][green]$};
\node  (N9) at (0,5) {$\Smiley[2][green]$};
\node (N10) at (2,5) {$\Smiley[2][green]$};
\node (N11) at (4,5) {$\Smiley[2][green]$};
\draw (foot) to [bend left=10] (N1);
\draw (foot) -- (N2);
\draw (foot) -- (N3);
\draw (foot) -- (N4);
\draw (foot) -- (N5);
\draw (foot) -- (N6);
\draw (foot) to [bend right=10] (N7);
\draw (foot) to [bend right=15] (N8);
\draw (N1) to [bend left=10] (N9);
\draw (N2) -- (N10);
\draw (N3) -- (N11);
\end{tikzpicture}
\caption{\label{fig:essai2}
The hydra $\iota(3,5)$}
\end{figure}

Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5]
\node (foot) at (2,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (3,4) {$\Smiley[2][green]$};
\node (N3) at (1,4) {$\Smiley[2][green]$};
\draw (foot) -- (N1);
\draw (N1) to [bend right =15] (N2);
\draw (N1) to  [bend left=15](N3);
\end{tikzpicture}
\caption{\label{fig:h-omega2-small}}
 The hydra \texttt{big\_h}.
\end{figure}
\begin{Coqsrc}
   Let big_h := hyd1 (hyd2 head head).  
 \end{Coqsrc}

 In a second step, we build a ``smaller'' hydra.

\begin{Coqsrc}
   Let small_h := iota (m big_h).
\end{Coqsrc}

Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h o<= m small\_h o< m big\_h}, which is impossible.

\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}

For proving the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
build a battle transforming \texttt{big\_h} into \texttt{small\_h}.

First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h}
into the pair $(i,j)$.
If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the folowing result.

\begin{Coqsrc}
 Lemma big_to_small: big_h -+-> small_h.
\end{Coqsrc}

Since $m$ is a variant, we infer the following inequality:

\begin{Coqsrc}
Corollary m_lt : m small_h o< m big_h.
\end{Coqsrc}

\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }

The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some pair $(i,j)$, where $i>0$, there exists an infinite number of
pairs stricly less than $(i,j)$, and there exists an infinite number of battles that start from
$\iota(i,j)$. In effect, at any configuration $\iota(k,0)$, the hydra can freely chose any replication number. Intuitively, the measure of such a hydra must be large enough for taking into account
all the possible battles issued from that hydra.
Let us give more technical details.

\begin{itemize}
\item The proof of the lemma \texttt{m\_ge : forall p : t,   p o<= m (iota p)} uses well-founded induction on $p$, and not structural induction on natural numbers

\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
  \begin{itemize}
  \item $p=(0,0)$
  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
  \end{itemize}
\end{itemize}

Before starting the proof, we have to express the notion of \emph{limit} in terms of
least upper bounds, through the following logical equivalence.

\vspace{4pt}
\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#limit_is_lub}{Hydra.Omega2\_Small.v}}.

\begin{Coqsrc}
Lemma limit_is_lub : forall i p, 
   (forall j, (i,j) o< p) <-> (S i, 0) o<= p.  
\end{Coqsrc}

Let us define the notion of elementary ``step'' of decreasing sequences in
\texttt{t}

\begin{Coqsrc}
Inductive step : t -> t -> Prop :=
| succ_step : forall i j,  step (i, S j) (i, j)
| limit_step : forall i j, step (S i, 0) (i, j).
\end{Coqsrc}

The following lemma establishes a correspondance between the relation
\texttt{step} and hydra battles.

\begin{Coqsrc}
Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
\end{Coqsrc}

\index{Maths!Transfinite induction}

Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
by transfinite induction over \texttt{p} a battle 
that transforms the hydra $\iota(p)$ into $\iota(q)$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
Lemma m_ge : forall p : t,   p o<= m (iota p).
Proof.
  intro p ; pattern p;
    apply  well_founded_induction with 
               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
     intros (i,j) IHij (* rest of proof skipped *)
\end{Coqsrc}

\begin{Coqanswer}
  i, j : nat
  IHij : forall y : t, y o< (i, j) -> y o<= m (iota y)
  ============================
  (i, j) o<= m (iota (i, j)) 
\end{Coqanswer}

Then we have  three cases to consider, according to the values of $i$ and $j$.
\begin{itemize}
\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
 remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
is any natural number.

Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.

Applying the induction hypothesis, we get the inequality
 $ m(\iota(i,j)) \geq (i,j)$ for any $j$.

Thus, $m(\iota(p)) > (i,j)$ for any $j$.
Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
$\iota(i+1,0)\geq (i+1,0)$

\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$

\end{itemize}

\subsubsection{End of the proof}

Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
variant for termination on the lexicographic square of $(\mathbb{N},<)$.

\begin{enumerate}
\item From \texttt{m\_lt}, we infer the strict inequality 
\texttt{m small\_h o< m big\_h}
\item from \texttt{m\_ge}, we get \texttt{m big\_h o<= m (iota (m big\_h)) = m small\_h} 
\end{enumerate}

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
Theorem Impossible : False.
Proof.
  destruct (StrictOrder_Irreflexive  (m big_h)).
  apply le2_lt2_trans with (m small_h).
  -  unfold small_h;  apply m_ge.
  -  apply m_lt. 
Qed.

\begin{exercise}
Prove that there exists no variant $m$ from \texttt{Hydra} into \texttt{nat*nat} for proving
    the  termination of all \emph{standard} battles.
\end{exercise}

\begin{remark}
In Chapter~\ref{ks-chapter}, we will prove a much more general  theorem
than the two previous results about $\mathbb{N}$ and $\mathbb{N}\times\mathbb{N}$. The proof of that general  result will share the same structure, but will 
require a lot of technical results. 
\end{remark}

\index{Exercises}
\begin{exercise}

\label{sec:orgheadline63}
Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.

\begin{itemize}
\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
  lexicographic ordering
\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.

For instance, the following inequality holds:
\[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
\end{itemize}

>>>>

%\index{To do}

\begin{todo}
Look for an existing library on trees with nodes of arbitrary arity, in order to replace  this ad-hoc type with something more generic.
\end{todo}

@@

### 2020-09-18 doc/chapter-ordinals.tex
update the index
modification in *.tex

\subsection{The ordinal \texorpdfstring{$\omega^2$}{omega^2}}

@@

@@

  - intros.  unfold m, measure_lt;  cbn; destruct xs0; simpl; left; lia.
  - intros;    unfold m, measure_lt ;cbn; destruct ys0; simpl; right; lia.
   - auto.
  Defined.

@@

@@

@@

\begin{Coqsrc}
Class  SubON 
       `(OA : @ON A ltA  compareA)
       `(OB : @ON B ltB  compareB)
       (alpha :  B)
       (iota : A -> B):=
  {
  subon_compare :forall x y : A,  compareB (iota x) (iota y) =
                                 compareA x y;
  subon_incl : forall x, ltB (iota x) alpha;
  subon_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Class ON_for `(alpha : Schutte_basics.Ord)
     `(OA : @ON A ltA  compareA)
      (iota : A -> Schutte_basics.Ord) :=
  { ON_for_inj : forall a, Schutte_basics.lt (iota a) alpha;
    ON_for_onto : forall beta, Schutte_basics.lt beta alpha ->
                                exists b, iota b = beta;
    On_compare_spec : forall a b:A,
        match compareA a b with
          Datatypes.Lt => Schutte_basics.lt (iota a) (iota b)
        | Datatypes.Eq => iota a = iota b
        | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
        end}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Instance Epsilon0_correct :
  ON_for epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
\end{Coqsrc}

@@

{\it \textbf{Note:s} This property reflects the equality $i+\omega=\omega$ we will prove in larger notation systems, as well as in Schütte's model}

>>>>

\section{The ordinal \texorpdfstring{$\omega^2$}{omega^2}}

@@

\index{Coq!Commands!Function}

@@

  - intros; unfold m, measure_lt; cbn; destruct xs0; simpl; left; lia.
  - intros; unfold m, measure_lt; cbn; destruct ys0; simpl; right; lia.
   - auto.
  Defined.

@@

\subsection{Yet another  proof of impossibility}
\label{omega2-case}

In Sect.~\vref{omega-case}, we proved that there exists no variant towards \texttt{nat}
(\emph{i.e.} the ordinal $\omega$) for proving the termination of all hydra battles.
We  prove now that even the ordinal $\omega^2$ is also insufficient for this purpose.

The proof we are going to develop has exactly the same structure as in Section~\ref{omega-case}.
 Nevertheless, the proof of technical  lemmas is a little more complex, due to 
 the structure of the lexicographic order on $\mathbb{N}\times\mathbb{N}$. 
Consider for instance that there exists an infinite number of ordinals  between
$\omega$ and $\omega\times 2$.

The detailed  proof script is in the file \url{../src/html/hydras.Hydra.Omega2_Small.html}.

\subsubsection{Preliminaries}
Let us assume there is a variant from \texttt{Hydra} into $\omega^2$  for proving the   termination of all hydra battles.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
  Section Impossibility_Proof.

 Variable m : Hydra -> ON_Omega2.t.

 Context (Hvar : Hvariant lt_wf free m).
\end{Coqsrc}

Let us follow the same pattern as in Sect.~\ref{omega-case}.
First, we define an injection from type \texttt{t} into \texttt{Hydra}, by
 associating to  each ordinal $\omega\times i+ j = (i,j)$ the hydra with $i$ branches of length $2$ and
$j$ branches of length $1$.

%% revenir ici

\vspace{4pt}
\emph{From Module ~\href{../src/html/hydras.Hydra.Omega2_Small.html\#iota}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
Let iota (p: ON_Omega2.t) := 
    node (hcons_mult (hyd1 head) (fst p)
                     (hcons_mult head (snd p) hnil)).
  \end{Coqsrc}

For instance, Figure~\vref{fig:essai2} shows the hydra associated to the ordinal 
$(3,5)$, a.k.a. $\omega\times 3 + 5$.

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.4]
\node (foot) at (6,0) {$\bullet$};
\node (N1) at (1,3) {$\bullet$};
\node (N2) at (3,3) {$\bullet$};
\node (N3) at (5,3) {$\bullet$};
\node (N4) at (8,3) {$\Smiley[2][green]$};
\node (N5) at (11,3) {$\Smiley[2][green]$};
\node (N6) at (14,3) {$\Smiley[2][green]$};
\node (N7) at (17,3){$\Smiley[2][green]$};
\node (N8) at (20,3){$\Smiley[2][green]$};
\node  (N9) at (0,5) {$\Smiley[2][green]$};
\node (N10) at (2,5) {$\Smiley[2][green]$};
\node (N11) at (4,5) {$\Smiley[2][green]$};
\draw (foot) to [bend left=10] (N1);
\draw (foot) -- (N2);
\draw (foot) -- (N3);
\draw (foot) -- (N4);
\draw (foot) -- (N5);
\draw (foot) -- (N6);
\draw (foot) to [bend right=10] (N7);
\draw (foot) to [bend right=15] (N8);
\draw (N1) to [bend left=10] (N9);
\draw (N2) -- (N10);
\draw (N3) -- (N11);
\end{tikzpicture}
\caption{\label{fig:essai2}
The hydra $\iota(\omega\times 3+5)$}
\end{figure}

Like in Sect.~\ref{omega-case}, we build a hydra out of the range of \texttt{iota} (represented in Fig.~\vref{fig:h-omega2-small}).

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5]
\node (foot) at (2,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (3,4) {$\Smiley[2][green]$};
\node (N3) at (1,4) {$\Smiley[2][green]$};
\draw (foot) -- (N1);
\draw (N1) to [bend right =15] (N2);
\draw (N1) to  [bend left=15](N3);
\end{tikzpicture}
\caption{\label{fig:h-omega2-small}}
 The hydra \texttt{big\_h}.
\end{figure}

\begin{Coqsrc}
   Let big_h := hyd1 (hyd2 head head).  
\end{Coqsrc}

 In a second step, we build a ``smaller'' hydra.

\begin{Coqsrc}
   Let small_h := iota (m big_h).
\end{Coqsrc}

Like in Sect.~\ref{omega-case}, we prove the double inequality \texttt{m big\_h o<= m small\_h o< m big\_h}, which is impossible.

\subsubsection{Proof of the inequality \texttt{m small\_h o< m big\_h}}

In order to prove the inequality  \texttt{m\_lt: m small\_h o< m big\_h}, it suffices to
build a battle transforming \texttt{big\_h} into \texttt{small\_h}.

First we prove that \texttt{small\_h} is reachable from \texttt{big\_h} in one or two steps. Let us decompose \texttt{m big\_h} as $(i,j)$.
If $j=0$, then one round suffices to transform \texttt{big\_h} into $\iota(i,j)$.
If $j>0$, then a first round transforms \texttt{big\_h} into $\iota(i+1,0)$ and a second round into $\iota(i,j)$. So, we have the following result.

\begin{Coqsrc}
Lemma big_to_small: big_h -+-> small_h.
\end{Coqsrc}

Since $m$ is a variant, we infer the following inequality:

\begin{Coqsrc}
Corollary m_lt : m small_h o< m big_h.
\end{Coqsrc}

\subsubsection{Proof of the inequality \texttt{m big\_h o<= m small\_h} }

The proof of the inequality \texttt{m big\_h o<= m small\_h} is quite more complex than in Sect~\ref{omega-case}.  If we consider some ordinal $\alpha=(i,j)$, where $i>0$, there exists an infinite number of
ordinals stricly less than $\alpha$, and there exists an infinite number of battles that start from
$\iota(\alpha)$. In effect, at any configuration $\iota(k,0)$, where $k>0$, the hydra can freely chose any replication number. Intuitively, the measure of such a hydra must be large enough for taking into account
all the possible battles issued from that hydra.
Let us give more technical details.

\begin{itemize}
\item The proof of the lemma \texttt{m\_ge : m big\_h  o<= m small\_h} uses well-founded induction on \texttt{big\_h}.

\item For any pair $p$, we have to distinguish between three cases, according to the value of $p$'s components.
  \begin{itemize}
  \item $p=(0,0)$
  \item $p=(i,0)$, where $i>0$\,: $p$ corresponds to a limit ordinal
  \item $p=(i,j)$, where $j>0$\,: $p$ is the successor of $(i,j-1)$.
  \end{itemize}
\end{itemize}

Let us define the notion of elementary ``step'' of decreasing sequences in
\texttt{t}

\begin{Coqsrc}
Inductive step : t -> t -> Prop :=
| succ_step : forall i j,  step (i, S j) (i, j)
| limit_step : forall i j, step (S i, 0) (i, j).
\end{Coqsrc}

The following lemma establishes a correspondance between the relation
\texttt{step} and hydra battles.

\begin{Coqsrc}
Lemma step_to_battle : forall p q, step p q -> iota p -+-> iota q.
\end{Coqsrc}

\index{Maths!Transfinite induction}

Thus, starting from any inequality $q < p$ on type \texttt{t}, we can build 
by transfinite induction over \texttt{p} a battle 
that transforms the hydra $\iota(p)$ into $\iota(q)$.

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html\#m_ge}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
Lemma m_ge : forall p : t,   p o<= m (iota p).
Proof.
  unfold small_h; pattern (m big_h) .   
   apply  well_founded_induction with (R := lt) (1:= lt_wf).
  intro p ; pattern p;
    apply  well_founded_induction with 
               (R := lt2) (1:= wf_lexico lt_wf lt_wf);
     intros (i,j) IHij. 
\end{Coqsrc}

\begin{Coqanswer}
  i, j : nat
  IHij : forall y : t, y o< (i, j) -> y o<= m (iota y)
  ============================
  (i, j) o<= m (iota (i, j)) 
\end{Coqanswer}

Then we have  three cases to consider, according to the values of $i$ and $j$.
\begin{itemize}
\item If $p=(0,0)$ then obviously, $\iota(p)\geq p = (0,0)$
\item If  $p=(i+1,0)$ for some $i\in\mathbb{N}$, we
 remark  that $p$ is strictly greater than any pair $ (i, j)$, where $j$ 
is any natural number.

Applying the battle rules, for any $j$, we have $\iota(i+1,j)  {\round} \iota(i, j) $, thus $m(\iota(p)) > m(\iota(i,j)$ since  $m$ is assumed to be a variant.

Applying the induction hypothesis, we get the inequality
 $ m(\iota(i,j)) \geq (i,j)$ for any $j$.

Thus, $m(\iota(p)) > (i,j)$ for any $j$.
Applying the lemma \texttt{limit\_is\_lub}, we get  the inequality
$\iota(i+1,0)\geq (i+1,0)$

\item If $p=(i,j+1)$ with $j\in\mathbb{N}$, we have  $\iota(p)  {\round} \iota(i, j) $,
hence $m(\iota(p))> m(\iota(i,j)) \geq (i,j)$, thus $m(\iota(p))\geq (i,j+1)=p$

\end{itemize}

\begin{Coqsrc}
  (* ... *)
Qed.
\end{Coqsrc}

\subsubsection{End of the proof}

Since $<$ is a strict order (irreflexive  and transitive) on \texttt{nat*nat}, we can conclude that there is no
variant for termination on the lexicographic square of $(\mathbb{N},<)$.

\begin{enumerate}
\item From \texttt{m\_lt}, we infer the strict inequality 
\texttt{m small\_h o< m big\_h}
\item from \texttt{m\_ge}, we get \texttt{m big\_h o<= m (iota (m big\_h)) = m small\_h} 
\end{enumerate}

\vspace{4pt}
\emph{From Module~\href{../src/html/hydras.Hydra.Omega2_Small.html}{Hydra.Omega2\_Small}}

\begin{Coqsrc}
Theorem Impossible : False.
Proof.
  destruct (StrictOrder_Irreflexive  (m big_h)).
  apply le2_lt2_trans with (m small_h).
  -  unfold small_h;  apply m_ge.
  -  apply m_lt. 
Qed.

End Impossibility_Proof.
\end{Coqsrc}

\begin{exercise}
Prove that there exists no variant $m$ from \texttt{Hydra} into $\omega^2$ for proving
    the  termination of all \emph{standard} battles.
\end{exercise}

\begin{remark}
In Chapter~\ref{ks-chapter}, we  prove a generalization of the impossibility lemmas of
Sect.~\ref{omega-case} and this section, with the same proof structure, but with much more 
complex technical details.
 \end{remark}

\index{Exercises}
\begin{exercise}

\label{sec:orgheadline63}
Write \emph{direct} proofs ({i.e.},  without applying the result and tools of Chap.~\ref{ks-chapter}) that the following data structures  are too simple for defining a variant for any hydra battle.

\begin{itemize}
\item  $\omega^n$ : the set of all $n$-uples of natural numbers, ordered  by 
  lexicographic ordering
\item  $\omega^\omega$: the set of all decreasing sequences (with respect to $\le$)  of natural numbers, ordered by lexicographic ordering on lists.

For instance, the following inequality holds:
\[\langle 4,3,3,3,3,3,3,2,2,2 \rangle\,<\,\langle 4,4,2 \rangle\]
\end{itemize}

@@

If \texttt{OB} is presumed to be correct, the we can consider that \texttt{OA} ``inherits'' its correcteness from the bigger notation system \texttt{OB}.

@@

\begin{Coqsrc}
Class  SubON 
       `(OA : @ON A ltA  compareA)
       `(OB : @ON B ltB  compareB)
       (alpha :  B)
       (iota : A -> B):=
  {
  subon_compare: forall x y : A,  compareB (iota x) (iota y) =
                                 compareA x y;
  subon_incl : forall x, ltB (iota x) alpha;
  subon_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Class ON_correct `(alpha : Schutte_basics.Ord)
     `(OA : @ON A ltA  compareA)
      (iota : A -> Schutte_basics.Ord) :=
  { ON_correct_inj : forall a, Schutte_basics.lt (iota a) alpha;
    ON_correct_onto : forall beta, Schutte_basics.lt beta alpha ->
                                exists b, iota b = beta;
    On_compare_spec : forall a b:A,
        match compareA a b with
          Datatypes.Lt => Schutte_basics.lt (iota a) (iota b)
        | Datatypes.Eq => iota a = iota b
        | Datatypes.Gt => Schutte_basics.lt (iota b) (iota a)
        end}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Instance Epsilon0_correct :
  ON_correct epsilon0 Epsilon0  (fun alpha => inject (cnf alpha)).
\end{Coqsrc}

@@

{\it \textbf{Note:} This property reflects the equality $i+\omega=\omega$ we will prove in larger notation systems, as well as in Schütte's model}
